\section{Literature Review}
\subsection{Introduction}
In Belgium one man in three and one woman in four faces cancer before his or
her 75th birthday \cite{kanker}. In 2010 62017 new cases of cancer were
diagnosed in Belgium \cite{kankerliga}. Lung cancer is the second most common
cander for men, and the third for women.
On top of that, it is one of the deadliest cancers
\cite{zheng}. However, an early detection can increase the survival rate up to 70-80\%
\cite{swensen}. Furthermore, research has shown that the detection of lung
cancer in an early stage broadens the amount of treatment options and increases
the amount of invasive surgery\cite{greenlee}.


Due to recent developments in
computed tomography (CT) technology it is now possible to obtain near isotropic, submillimeter resolution images of the complete chest in a single breath hold.
This high resolution has the advantage that it enables visualisation of small
and low-contrast nodules that could hardly be screened in conventional
programs. The downside is that enormous amounts of data are generated
which increases the work load of radiologists, especially since low-dose CT
scans are more and more implemented in routine screenings. Still, this is no
idle measure. Long nodules are very commonly detected on CT scans. Research
shows that up to 51\% of smokers aged 50 years or older have pulmonary lung nodules on CT scans \cite{mahon}.
Therefore, the United States Preventive Services Task Force for example stated that it ``recommends annual screening for lung cancer with low-dose computed tomography (LDCT) in adults aged 55 to 80 years who have a 30 pack-year smoking history and currently smoke or have quit
within the past 15 years'' \cite{ups}.
Therefore, the detection of pulmonary nodules from volumetric computed
tomography (CT) scans if one of the most studied CAD applications
\cite{sluimer}.


Currently, expert radiologists perform the investigation of the
CT scans. They use the shape, the texture, the location and the growth rate of
the volume of the nodule as clinical parameter to determine the malignancy of
the nodules and to decide on the diagnosis of lung cancer. A jagged shape
nodule is more likely to be lung cancer than a smooth one. A fatty, bony,
watery nodule or a mixture of these different contents is less likely to
indicate lung cancer than a nodule that is attached to a vessel. A nodule
attached to the lung wall is typically diagnosed as benign if the
volume-doubling periode is longer than 400 days \cite{wu}. Nevertheless, the examination of these scans is a time-consuming task and is not free from errors. Although small nodules are
in principle detectable in CT scans, a non-negligible fraction may be overlooked
if they are situated in a maze of vessels of similar size \cite{ozekes}.
FIGUUR
Another problem that arises is the intra- and interreader variability
amongst radiologists in pulmonary lung detection \cite{armato} \cite{hens}. Therefore,
there is a need for a computer-aided detection (CAD) system that can assist the
radiologist in the detection of pulmonary nodules.

\subsection{The biology of lung nodules}
FIGUUR vanginneken
Lung nodules are lung tissue abnormalities that are roughly spherical with a
diameter up to 30 mm. On chest CT scans they appear as a rounded or
irregular opacity. Many types of lung nodules can by distinguished on CT scans.
A centrilobular nodule is separated by several millimeters from the pleural surfaces, fissures and
interlobular septa. They range in size from a few millimeters to 10 millimeters.
A micronodule is less than 3 millimeters in diameter. A ground-glass nodule --
or non-solid nodule -- appears on the CT scans as a hazy attenuation in the
lung. This type of nodule does not efface the bronchial and vascular margins. A solid
nodule shows a homogeneous soft-tissue attenuation. Finally, a part-solid nodule
exhibits both ground-glass and solid soft-tissue attenuation characteristics
\cite{nodule}. 

The types of nodules stated above can be categorised. Juxta-vascular
pulmonary nodules have significant connections to their neighbouring vessels.
Pleural tail nodules have only thin connections to the neighbouring pleural
wall. Well-circumscribed nodules on the other hand do not have a connection to
the neighbouring vessels and structures. Juxta-pleural nodules show some degree
of attachment to their neighbouring pleural surface \cite{kostis}.

A number of nodule segmentation algorithms perform well in detecting specific
types of nodules e.g. large, spherical, isolated nodules. However, these CAD
systems show large limitations in detecting e.g. non-isolated nodules that are
connected to the pulmonary wall \cite{keshani}. These algorithms can be usefull
in particular situations, but if a detection algorithms really aims at being an
asset for the radiologist, it should be able to detect all nodules while
refusing as much false positives as possible.

\subsection{Overview of existing lung nodule detection systems}
As the demand for a reliable CAD system to detect pulmonary nodules is urgent, a
lot of research has been dedicated to the matter. Several commercial systems
have already been developed and many workstations that radiologists use to
examinate CT scans offer on-board nodule detection or enhancement capabilities
\cite{ginneken}. However, although a lot of efforts were made, the results shown
in the various studies are rather diverse.

\subsubsection{Commercial systems}
In 2004 iCAD, Inc., provided lung cancer detection, analysis and tracking
software for the TeraRecon's Aquarius product line. The latter licensed three
software modules from iCAD. The iCAD QuickCueTM for example automatically
detects cancerous lung nodules while the iCAD QuickMatchTM locates, compares and tracks these
nodules in previous or subsequent patient studies \cite{tera}.
However, the ImageChecker CT, launched by R2 Technology, was the first CT Lung
CAD system approved by the US Food and Drug Administration for the detection of
lung nodules during the examination of CT scans \cite{Mevis}. In 2005 R2
Technology, Inc., introduced the second-generation ImageChecker CT Lung Version 2.0 CAD system which also implemented the AutoPoint temporal comparison algorithm. This CAD system ``highlights abnormalities'' and compares new and past images to
demonstrate changes that have occured over time \cite{diag} \cite{r2}.
In 2006 Vital Images, Inc., and R2 Technology, Inc., announced the
implementation of the R2 Technology's ImageChecker CT Lung CAD software into
the Vitrea workstations \cite{vital}. In 2006 anohter company, Hologic, Inc.,
acquired R2 Technology, Inc., and implemented their CAD technology
\cite{Hologic}. Then, in 2008, MeVis Medical Solutions AG, Inc., acquired the
Pulmonary Computed Tomography Business from Hologic R2, Inc. \cite{Mevis}. 


Although the technology from R2 Technology is highly demanded, some companies
have also developed their own software. In 2007 Medicsight plc announced it was
granted a medical device license from the Therapeutic Products Directorate of
Health in Canada to introduce Medicsight LungCAD API \cite{HI}.
Median Technologies offers the LMS-Lung and LMS-Lung/CAD modules which provide
quantification and detection functionalities for pulmonary (solid) nodules and
micronodules \cite{median}. And Siemens has developed the syngo.CT Lung CAD.
They claim it is ``a fully automated computer assisted second reader tool'' that is designed to assist radiologists in
the detection of solid pulmonary nodules \cite{siemens}.

\subsubsection{Automated lung detection systems: publications}
Apart from these numerous commercial systems also a lot of academic research
centra have tried to come up with a successful pulmonary nodule detection
system. \cite{review} suggest that most CAD systems for the automated detection
of lung nodules proceed according to a number of steps of which the first one is
the acquisition of data. The detection of lung nodules is preferably performed
in CT scans as they enable the visualisation of small volume and low-contrast
nodules because of the limited slice thichness. A large number of chest CT scans
are available in public databases such as the LIDC database. During the next
step, the data are pre-processed to remove noise and artefacts. This might
improve the quality of the images, but it is not necessary to do so. In the
third step a segmentation of the lungs is performed. The lung lobe region is
identified and the rest of the image is removed. This reduces the computational
cost compared to the case where the whole image is processed and it increases
the reliability, the accuracy and the precision of the algorithm. This
increases the performance of the next step: the nodule detection. ``Lung nodule
detection refers to the process of determining whether nodule patterns are
present in the image and identifying the location of the nodules''
\cite[p.~154]{review}. Nodule detection methods can be categorized according to
the detection method that is applied. The first group of publications uses a
classification technique to classify voxels or regions of interest (ROI). In
addition, a clustering method may be implemented to improve the perfromance of
the automated nodule detection method. The second group uses template matching
to detect specific geometries. The third group relies entirely on the output of
a lung nodule segmentation method. The systems that include a classification
component in their nodule detection algorithm have demonstrated better
performances \cite{review}. In the final step, the amount of false positives is
reduced to achieve a maximum sensitivity. In the following paragraphs a summary
of relevant literature is given.

\cite{elbaz} proposed an three step algorithm to isolate lung nodules from
spiral chest low-dose CT (LDCT) scans. In the first step the lung tissues were
isolated by applying an iterative Markov-Gibbs-random-field (MGRF) based
segmentation framework. To retain the nodules attached to the pulmonary walls, the
segmentation was refined by the iterative conditional mode relaxation that
maximizes a MGRF energy. Then the lung nodules, arteries, veins, bronchi and
bronchioles were separated from the rest of the tissues in the slice. In the
second step the lung nodules (2-12 mm) were detected by applying 3D and 2D
templates which describe typical geometry and greylevel distributions within
nodules of the same type. Four template shapes were used: solid sphere, hollow
sphere, solid circle and solid semicircle. The radius and the greyscale
intensity of the templates was made adaptable. The detection combined the
normalized cross-correlation template matching and a genetic optimization
algorithm. The third step eliminated the false positive nodules using three
textural and geometric features that were calculated for each detected nodule.
To distuingish between false positives and true positives Bayesian supervised
classifier learning statistical characteristics from a training set (20 FP, 20
lung TP, 20 lung wall TP nodules) of nodules selected from 50 separate subjects.
CT scans from 200 subjects were used in this study. The sensitivity was 82,3\%
and a FPNs rate of 9,2\% (i.e. the number of FPNs -- 12-- with respect to the
total number of true nodules -- 130 -- ). The speed of the execution is a
function of the CPU and the data size. The algorithm, that was implemented in
C++ on an Intel dual processor wiht 16 GB memory and 2 TB hard drive, took about
5 minutes to process 182 CT slices of size 512 x 512.



segmentation

classification
	SVM
	RF
	ANN
	
ANODE

\subsubsection{Performance of existing systems}
 The algorithms presented in a wide range of papers report varying
 successes in the automated detection of nodules. However, it is very difficult
 to compare studies against one another in a meaningful way due to differences
 in the size of the datasets, the evaluation methods, the data selection
 criteria and the characteristics of the nodules \cite{lee2010}. Especially comparing older and contemporary studies is
 difficult as older ones may have used scans with thicker sections (range 2.5 -
 10 mm), on which small nodules are rather difficult to detect, than the scans
 nowadays (2,5 mm) \cite{lee2010, ginneken} \cite{mur}. Some studies
 focus on nodules below or above a certain size or on special types of nodules
 (e.g. solid nodules). \cite{mur} performed an extended
 literature review and found that the number of scans used for testing varied between 5 and 500 with a
 median number of 29,5. Many of the studies included multiple scans from
 individual patients, which means that the diversity of the available nodules
 was reduced. Furthermore, the results of publications are often presented in a
 diverse way. \cite{results} suggests the performance of CAD systems should be
 presented a the sensitivity of the system, the specificity, the accuracy, the
 error rate, the True Positive Rate and the False Positive Rate.
 
 In order to improve the access to data, and hereby the comparability between
 studies, the Lung Image Data Consortium (LIDC) created a publically available
 database which provides researchers with a vast amount of test- and
 trainingsdata. Nevertheless, as one can take different subdatasets from this
 large database for the training and the testing of algorithms, it is still not
 possible to compare results in an objective and meaningful way. Therefore,
 \cite{ginneken} created ANODE09, a database of 55 scans and a web-based
 framework which allows researchers to test their algorithms and to
 compare results against one another.
 
\subsection{Ensemble classifiers}
- reden om classifiers te gaan gebruiken

\subsubsection{Introduction to ensemble classifiers}
voordelen/nadelen verschillende classifiers

waarom random forest

\subsubsection{Random Forests}
wisk achtergrond







